---
title: "How to use CSICalc"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  # echo = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

To use this package, you need to install it first. You can install the development version from GitHub with the following two commands:

``` r
install.packages("devtools")
devtools::install_github("dmorrison01/CSICalc")
```

Now, here's an example of how to use this package.

```{r, include = FALSE}


N0 = 15845
N0_formatted = format(N0, big.mark = ",")
X = c(58+11)
N1 = 4376+219
mu_hat = c(130)
mu_hat_in_years = round(mu_hat/365,2)
mu_CI_low = c(118)
mu_CI_high = c(142)
est.incidence = X/(mu_hat/365)/N0 * 100

```

Consider the MAA "LAg < 1.5 OD-n, Viral Load > 1000 copies/mL". The estimated mean window period from Duong et al (2015) was 
$\hat{\mu} = `r mu_hat`$ days (95% CI `r mu_CI_low`, `r mu_CI_high`); i.e., `r mu_hat_in_years` years.

In the HPTN 071 PC24 cross-section, there were $N_1 = `r N1`$ seropositive individuals, and $X = `r X`$ of these individuals were in the window period for this MAA.
(One individual did not have a viral load result; for now, let us assume that this individual would have been classified as MAA-negative). 
There were $N_0 = `r N0_formatted`$ seronegative individuals.

So we estimate incidence as:
$$\hat{\iota} = \frac{X}{N_0 \times \hat{\mu}} \times 100\% = \frac{`r X`\text{ new seroconversions}}{`r N0_formatted`\text{ persons at-risk} \times `r mu_hat_in_years`\text{ years}} \times 100\%$$
$$= `r round(est.incidence, 2)`\text{ new seroconversions per 100 person-years}$$

To compute a confidence interval for this incidence estimate, we need to account for the uncertainty in $\hat{\mu}$.
To do so, we assume a $\text{gamma}(\alpha, \beta)$ prior distribution for $\mu$, with probability density function
$$p(\mu) = \frac{1}{\beta^\gamma\Gamma(\gamma)} \mu^{\gamma-1}\exp\{-\mu / \beta\}$$

such that $E(\mu) = \gamma\beta = \hat{\mu} = `r mu_hat`\text{ days}$ and $P(\mu \in [`r mu_CI_low`, `r mu_CI_high`]) \approx 0.95$.

We can find the prior parameters $\gamma$ and $\beta$ using the `fit_gamma_distribution()` function:

```{r}
library(CSICalc)
prior = fit_gamma_distribution(
  mean = 130,
  lower = 118,
  upper = 142,
  confidence_level = .95,
  gamma_shape_range = c(0,1000))

print(prior)
```

We can check that those parameters produce the specified coverage using the following calculation:

```{r, include = TRUE}

pr1 = pgamma(shape = prior[1], scale = prior[2], q = 142) - 
  pgamma(shape = prior[1], scale = prior[2], q = 118)
   

print(pr1)
```

In other words, if $\gamma = `r round(prior[1], 2)`$ and $\beta = `r round(prior[2],2)`$, then
$P(\mu \in [`r mu_CI_low`, `r mu_CI_high`]) = `r round(pr1, 3)`$.

Now, let $I$ denote the incidence rate. Then $E[X] \approx N_0I\mu$ (Brookmeyer 1997). If $I\mu$ is small, then 
the conditional distribution of $X$, given $\mu$, will be approximately a Poisson distribution,
and the marginal distribution of $X$ (not conditional on $\mu$) will be a negative binomial distribution; specifically,
if $\mu \sim \text{gamma}(\gamma, \beta)$, then

$$P(X=x) \approx \frac{\Gamma(\gamma + x - 1)}{\Gamma(\gamma)x!} \left(\frac{\beta N_0 I}{\beta N_0 I + 1}\right)^x \left(\frac{1}{\beta N_0 I + 1}\right)^\gamma$$

We can compute a 95% confidence interval for the incidence rate, $I$, by finding $(i_1,i_2)$ such that:
$P(X\ge x| I = i_1) = 2.5\%$ and $P(X\le x| I = i_2) = 2.5\%$.
The following function computes that confidence interval:

```{r}

est_and_CI = est.inc(
    X = 69,
    N0 = 15845,
    mu_estimate = 130,
    mu_shape = prior[1],
    mu_scale = prior[2],
    incidence_CI_level = .95)

print(est_and_CI * 100 * 365) # rescale to get infections per 100 person-years

```

```{r, include = FALSE}

ci1 = round(est_and_CI[-1] * 100 * 365, 2)

```

Thus, given the data described above, our 95% confidence interval for the incidence rate is $(`r ci1[1]`, `r ci1[2]`)$ new infections per 100 person-years.

To do everything in one step, input the confidence interval for $\mu$ into `est.inc()`, like so:

```{r}

est_and_CI_v2 = est.inc(
    X = 69,
    N0 = 15845,
    mu_estimate = 130,
    mu_CI_low = 118,
    mu_CI_high = 142,
    mu_CI_level = .95,
    incidence_CI_level = .95)

print(est_and_CI_v2 * 100 * 365) # rescale to get infections per 100 person-years

```

## References:

Brookmeyer R. Accounting for follow-up bias in estimation of human immunodeficiency virus incidence rates. JRSS:A 1997; 160:127-140.

Duong YT, Kassanjee R, Welte A, Morgan M, De A, Dobbs T, et al. Recalibration of the limiting antigen avidity EIA to determine mean duration of recent infection in divergent HIV-1 subtypes. PLoS One 2015; 10(2):e0114947.

